Marlowe: implementing and analysing financial contracts\thanks{Supported by IOHK.}

### 2019-12-03 papers/marlowe2019/paper.tex
changed continuation type layout
Fix some typos and address SJT comments

For Turing-complete languages, this process is usually undecidable in the general case, and for Marlowe contracts it is often NP-complete in general. However, in practice, SAT and SMT (Satisfaction Modulo Theories) solvers are often able to solve moderately large instances of this problem reasonably efficiently.

Our implementation relies on the Haskell library SBV, which in turn relies on existing SAT and SMT solvers to check satisfiability of properties or finding counter-examples.

@@

SBV (SMT Base Verification) library provides a high-level API that allows developers to automatically prove properties about Haskell programs, among other functionalities. 
The SBV library %seamlessly 
translates these properties to SMTLib queries, passes them to one or several SAT/SMT solvers, and translates the results back to the format in which the queries were written.

@@

\noindent
can be used both with normal integers and with symbolic parameters:
%SJT: would be usweful to say something more about what symbolic parameters mean here: it's essentially that these are parameters that are universally quantified and passed to the solver, I think? They are instantiated with counter-examples if those exist. 
\begin{verbatim}
*Main> primeFactors 98 314
True
*Main> prove primeFactors
Falsifiable. Counter-example:
    s0 = 6353 :: Integer
    s1 = 4049 :: Integer
*Main> primeFactors 6353 4049
False
\end{verbatim}

@@

Secondly, to the best our knowledge, SBV does not currently provide support for custom data types directly, except for zero-ary datatypes, i.e: enumerations.
%SJT: i.e. there's no symbolic representation of user defined data types in general?

@@

We can see that this version of the implementation takes two concrete parameters: 
%SJT: need to explain a little more about what the bounds are
\texttt{Bounds} (with info about the bounds of the contract) and \texttt{Contract} (the contract we want to analyse); and two symbolic parameters: \texttt{SSlotNumber} (the initial slot number) and \texttt{SList NTransaction} (the input list of transactions); and it returns \texttt{SList NTransactionWarning} (a symbolic list of transaction warnings).

@@

\noindent
Because working with nested types is much more error prone than working with the original data-types, we used Template Haskell % SJT: nice! Need to add reference.
to implement functions that transform the custom datatypes into nested types and generate the appropriate conversion functions.

@@

In the non-symbolic semantics, each step typically takes a \texttt{Contract}, a \texttt{State}, and a set of \texttt{Input}s; and it returns an updated \texttt{State} and a remaining \texttt{Contract}. 
%SJT: a thought: could just give the type here, or cross-refer to an earlier section.
This is a problem because values that rely on symbolic values have to be themselves symbolic. Moreover,  the resulting remaining \texttt{Contract} depends on the \texttt{Input}s and \texttt{State}, which are both symbolic.

@@

However, in both the input and the \texttt{State} record there are several lists (representing finite maps) that are not explicitly bounded in the implementation. Some parts of the semantics are bounded by the length of these lists (or maps), such as the implementation of \texttt{Close} that refunds any remaining money in the contract accounts.
In order for the symbolic implementation to be finite we need to find a bound for the length of these lists and maps.

With the current semantics of Marlowe all the lists/maps in \texttt{State} and input can be bound when looking at a particular contract. In particular, we can find out quite easily, by looking at the contract, how many parties there are, how many choice identifiers are used, how many accounts are used, and how many let binding identifiers are used.

It is less obvious that we can also set a bound for the input \texttt{Transaction} list as the maximum number of nested \texttt{When}s in the contract (for all the execution paths) plus one. We also provide a proof in Isabelle that shows that any longer list of transactions will result in an error. 
%SJT: have I rephrased this right?
That is because any transactions that result in errors do not modify the \texttt{State} or the \texttt{Contract}, we know that limiting the analysis to lists of transactions that are shorter than the bound does not lose generality.

@@

>>>>

For Turing-complete languages, this process is usually undecidable in the general case, and for Marlowe contracts it is often NP-complete in general. However, in practice, SMT (Satisfaction Modulo Theories) solvers are often able to solve moderately large instances of this problem reasonably efficiently.

Our implementation relies on the Haskell library SBV, which in turn relies on existing SMT solvers to check satisfiability of properties or finding counter-examples.

@@

SBV \cite{SBV} (SMT Base Verification) library provides a high-level API that allows developers to automatically prove properties about Haskell programs, among other functionalities. 
The SBV library translates these properties to SMTLib queries, passes them to one or several SMT solvers, and translates the results back to the format in which the queries were written.

@@

\noindent
can be used both with normal integers and with symbolic parameters, that is, parameters whose concrete value is not specified and which the solver will try to replace with values that satisfy or falsify the property:

\begin{verbatim}
*Main> primeFactors 98 314
True
*Main> prove primeFactors
Falsifiable. Counter-example:
    s0 = 6353 :: Integer
    s1 = 4049 :: Integer
*Main> primeFactors 6353 4049
False
\end{verbatim}

@@

Secondly, to the best our knowledge, SBV does not currently directly provide support for representing custom data types symbolically, except for zero-ary datatypes, i.e: enumerations.

@@

We can see that this version of the implementation takes two concrete parameters: \texttt{Bounds} (see \textit{Bounds for the state and the inputs} section below) and \texttt{Contract} (the contract we want to analyse); and two symbolic parameters: \texttt{SSlotNumber} (the initial slot number) and \texttt{SList NTransaction} (the input list of transactions); and it returns \texttt{SList NTransactionWarning} (a symbolic list of transaction warnings).

@@

\noindent
Because working with nested types is much more error prone than working with the original data-types, we used Template Haskell \cite{sheard2002template} to implement functions that transform the custom datatypes into nested types and generate the appropriate conversion functions.

@@

In the non-symbolic semantics, each step typically takes a \texttt{Contract}, a \texttt{State}, and a set of \texttt{Input}s; and it returns an updated \texttt{State} and a remaining \texttt{Contract}.

This is a problem because values that rely on symbolic values have to be themselves symbolic. Moreover,  the resulting remaining \texttt{Contract} depends on the \texttt{Input}s and \texttt{State}, which are both symbolic.

@@

However, in both the input and the \texttt{State} record there are several lists (representing finite maps) that are not explicitly bounded in the implementation. Some parts of the semantics are bounded by the length of these lists (or maps), such as the implementation of \texttt{Close} that refunds any remaining money in the contract accounts.
In order for the symbolic implementation to be finite, we need to find a bound for the length of these lists and maps.

With the current semantics of Marlowe, all the lists/maps in \texttt{State} and input can be bounded when looking at a particular contract. In particular, we can find out quite easily, by looking at the contract, how many parties there are, how many choice identifiers are used, how many accounts are used, and how many let binding identifiers are used.

It is less obvious that we can also set a bound for the input \texttt{Transaction} list as the maximum number of nested \texttt{When}s in the contract (for all the execution paths) plus one. We also provide a proof in Isabelle that shows that any longer list of transactions will result in an error (we discuss this in Section~\ref{subsubsec:bound_max_transaction_number}).

@@

\subsubsection{Bound on the maximum number of transactions\label{subsubsec:bound_max_transaction_number}}

\emph{ToDo}
Any transactions that result in errors do not modify the \texttt{State} or the \texttt{Contract}, we know that limiting the analysis to lists of transactions that are shorter than the bound does not lose generality.
